Synthesizing Symmetric Lenses

### 2018-07-10 papers/SymmetricLenses/main.tex
some comments / changes to section 4

\section{Simple Symmetric String Lenses}
\label{sec:ssl}
We now give a formal description of all the combinators in our symmetric lens
language. These lens terms are typed between pairs of stochastic regular
expressions: the typing judgment $\Lens \OfType \BRegex \Leftrightarrow
\BRegexAlt$ means that $\Lens$ is a well typed lens between
$\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$. While we are performing
synthesis on stochastic regular expressions, lenses are typed between regular
expressions.  We can recover a regular expression from a stochastic regular
expression by deleting probability annotations.  In a slight abuse of notation,
if $\Regex$ is a stochastic regular expression, $\BRegex$ is the regular
expression obtained by removing probability annotations.

@@

    $\REntropyOf{\Regex \Given \InvertOf{\Lens}, \RegexAlt}$
    & =
    & $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$\\
  \end{tabular}
\end{center}
$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ is defined symmetrically. These
functions bound the expected number of bits to recover one data format from a
synchronized string in the other format. Notice, we would be able to exactly
calculate the conditional entropy, were it not for \MergeL. If
$\MergeLOf{\Lens_1}{\Lens_2} \OfType \Regex \Leftrightarrow
\PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}$, given a string in $s$, we
need to determine if the synchronized string is in $\Regex$ or $\RegexAlt$.
However, this information content is dependent on the how likely the
synchronized string is to be in $\Regex$ or $\RegexAlt$. Furthermore, we
typically calculate the conditional entropy exactly, as merges are relatively
uncommon on practice; only \afm{2} of the lenses sythesized in our benchmarks
include merges.

@@

\begin{theorem}
  Let $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, where $\Lens$ does not
  include composition, $\Regex$ and $\RegexAlt$ and unambiguous, and neither
  $\Regex$ nor $\RegexAlt$ contain any empty subcomponents.
  \begin{enumerate}
  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ bounds the entropy of
    $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$, given $\SetOf{s
      \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd \Lens.\PutROf{s}{t} = t}$
  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ bounds the entropy of
    $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$, given $\SetOf{t \SuchThat t
      \in \LanguageOf{\RegexAlt} \BooleanAnd \Lens.\PutLOf{t}{s} = s}$
  \end{enumerate}
\end{theorem}

Note we have not defined entropy for lens compositions. Consider the lens
$\Lens_1 ; \Lens_2$, where both $\Lens_1$ and $\Lens_2$ are non-bijective.
However, their composition could actually be bijective, for example if the
intermediary format contains information not present in either the source nor
the target. While this composition may be a bijection (and should have zero
cost), it is difficult to see this looking merely at the syntax.

Notice that our definition of $\Entropy$ contains no case for sequential
composition $\Lens_1 ; \Lens_2$ and our theorem excludes lenses that contain
such compositions. Unfortunately, defining the entropy of lenses involving
composition is challenging because $\Lens_1$ might, for instance, add some
information that is subsequently projected away in $\Lens_2$. Such operations
can cancel, leaving us with a zero-entropy bijection. However, detecting such
cancellations directly is complicated. Fortunately, we are able to avoid such
complications by synthesizing \emph{DNF lenses} — simple symmetric lenses that
inhabit a disjunctive normal form. DNF lenses have a number of properties that
facilitate synthesis: they cut down the program search space dramatically and
they do not contain instances of composition.

>>>>

\section{Simple Symmetric String Lenses}
\label{sec:ssl}
\saz{This paragraph used to say that lenses are typed between stochastic REs
  and it also says that lenses are typed between REs.  I think its should just
  be REs? Double check that this is correct}

We now give a formal description of all the combinators in our symmetric lens
language.  While our algorithms perform synthesis on stochastic regular
expressions, lenses themselves are typed between regular expressions.  We can
recover a regular expression from a stochastic regular expression by deleting
probability annotations.  In a slight abuse of notation\saz{Why is this an
  abuse?}, if $\Regex$ is a stochastic regular expression, $\BRegex$ is the
regular expression obtained by removing probability annotations. The typing
judgment $\Lens \OfType \BRegex \Leftrightarrow \BRegexAlt$ means that $\Lens$
is a well typed lens between $\LanguageOf{\BRegex}$ and
$\LanguageOf{\BRegexAlt}$.

@@

    $\REntropyOf{\Regex \Given \InvertOf{\Lens}, \RegexAlt}$
    & =
    & $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$\\
  \end{tabular}
\end{center}
$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ is defined symmetrically. These
functions bound the expected number of bits to recover one data format from a
synchronized string in the other format. Note that we would be able to exactly
calculate the conditional entropy, were it not for \MergeL. If
$\MergeLOf{\Lens_1}{\Lens_2} \OfType \Regex \Leftrightarrow
\PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}$, given a string in $s$, we
need to determine if the synchronized string is in $\RegexAlt_1$ or $\RegexAlt_2$.
However, this information content is dependent on the how likely the
synchronized string is to be in $\RegexAlt_1$ or $\RegexAlt_2$. Nevertheless, we
typically calculate the conditional entropy exactly, as merges are relatively
uncommon on practice; only \afm{2} of the lenses sythesized in our benchmarks
include merges.

@@

\begin{theorem}
  Let $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, where $\Lens$ does not
  include composition, $\Regex$ and $\RegexAlt$ are unambiguous, and neither
  $\Regex$ nor $\RegexAlt$ contain any empty subcomponents.
  \begin{enumerate}
  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ bounds the entropy of
    $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$, given $\SetOf{s
      \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd \Lens.\PutROf{s}{t} = t}$
  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ bounds the entropy of
    $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$, given $\SetOf{t \SuchThat t
      \in \LanguageOf{\RegexAlt} \BooleanAnd \Lens.\PutLOf{t}{s} = s}$
  \end{enumerate}
\end{theorem}

% SAZ: This looks like a bad merge, since the paragraph below is better
% Note we have not defined entropy for lens compositions. Consider the lens
% $\Lens_1 ; \Lens_2$, where both $\Lens_1$ and $\Lens_2$ are non-bijective.
% However, their composition could actually be bijective, for example if the
% intermediary format contains information not present in either the source nor
% the target. While this composition may be a bijection (and should have zero
% cost), it is difficult to see this looking merely at the syntax.

Note that our definition of $\Entropy$ contains no case for sequential
composition $\Lens_1 ; \Lens_2$ and our theorem excludes lenses that contain
such compositions. Unfortunately, defining the entropy of lenses involving
composition is challenging because $\Lens_1$ might, for instance, add some
information that is subsequently projected away in $\Lens_2$. Such operations
can cancel, leaving a zero-entropy bijection composed from two non-zero entropy
transformations. However, detecting such cancellations directly is complicated and
this property is difficult to determine merely from syntax. Fortunately, we are able to
avoid such complication altogether by synthesizing \emph{DNF lenses}---simple symmetric
lenses that inhabit a disjunctive normal form. DNF lenses have a number of
properties that facilitate synthesis: they cut down the program search space
dramatically and they do not contain instances of composition.
